---
title: "EMA intervention sensitivity analyses"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 300,
                      colormodel = "cmyk", fig.width=7, fig.height=5)
options(scipen = 999)
```

This code reproduces the EMA intervention sensitivity analyses reported in the following manuscript:

[Using neural signatures to predict health behavior change: A mindful distancing intervention to reduce alcohol consumption as a proof of concept]()

# load packages
```{r}
library(pacman)
pacman::p_load(tidyverse, brms, ggeffects, kableExtra, tidybayes, install = TRUE)
```

# define aesthetics
```{r}
palette = c("#e64626", "#1985a1", "#4c5c68", "#FAC748")

plot_aes = theme_minimal() +
  theme(legend.position = "top",
        legend.text = element_text(size = 16),
        plot.title = element_text(hjust = 0.5),
        text = element_text(size = 18, family = "Futura Medium"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(color = "black"),
        axis.line = element_line(colour = "black"),
        axis.ticks.y = element_blank())
```

# load data
```{r}
merged = read.csv("../data/task_neuro_data.csv", stringsAsFactors = FALSE)
ema = read.csv("../data/ema.csv")
maas = read.csv("../data/MAAS.csv", stringsAsFactors = FALSE)
ratings = read.csv("../data/ratings.csv", stringsAsFactors = FALSE)
```

## merge and prep for modeling
```{r}
between = merged %>%
  filter(condition == "mindful distancing") %>%
  select(pID, dot, trial_cond, condition) %>%
  group_by(pID, trial_cond, condition) %>%
  summarize(dot_between = mean(dot, na.rm = TRUE)) %>%
  group_by(trial_cond) %>%
  mutate(dot_between_c = scale(dot_between, scale = FALSE, center = TRUE)) %>%
  mutate(sd_dot = sd(dot_between, na.rm = TRUE),
         dot_between_std = dot_between_c / sd_dot) %>%
  select(pID, condition, trial_cond, dot_between_std) %>%
  mutate(trial_cond = sprintf("%s_expression", trial_cond)) %>%
  spread(trial_cond, dot_between_std)

ema_within = ema %>%
  left_join(., between) %>%
  left_join(., maas) %>%
  left_join(., ratings)
```

# effectiveness and individual differences: experience sampling intervention analyses (H3-4) {.tabset}
## run model
```{r, cache = FALSE, echo = TRUE, results = "hide"}
prior = c(prior(normal(0, 1), class=b))

fm = bf(mindful_response ~ active_week * regulation_expression + 
                           active_week * confidence_rating + 
                           active_week * MAAS_mean + 
          (0 + active_week |i| pID)) +
  bf(craving_previous ~ active_week * regulation_expression + mindful_response * regulation_expression + 
                        active_week * confidence_rating + mindful_response * confidence_rating + 
                        active_week * MAAS_mean + mindful_response * MAAS_mean + 
       (0 + active_week + mindful_response |i| pID)) +
  bf(drinks_number ~ active_week * regulation_expression + regulation_expression * craving_previous + mindful_response * regulation_expression  + 
                     active_week * confidence_rating + confidence_rating * craving_previous + mindful_response * confidence_rating  + 
                     active_week * MAAS_mean + MAAS_mean * craving_previous + mindful_response * MAAS_mean  + 
       (0 + active_week + craving_previous + mindful_response |i| pID)) +
  set_rescor(FALSE)
          
fit_brm_sequential = brm(
  fm,
  data  = ema_within,
  cores = 4,
  thin  = 4,
  seed  = 6523,
  control = list(adapt_delta = .99, max_treedepth = 15),
  prior = prior
)
```

## table
```{r}
fit_brm_sequential %>%
  broom.mixed::tidy(conf.int = TRUE, conf.level = 0.95) %>%
  filter(effect == "fixed") %>%
  mutate(term = gsub("\\(Intercept\\)", "intercept", term),
         term = gsub("regulation_expression", "signature expression", term),
         term = gsub("active_weekon", "intervention week (active)", term),
         term = gsub("active_weekoff", "intervention week (control)", term),
         term = gsub("mindful_response", "mindful response", term),
         term = gsub("craving_previous", "craving previous", term),
         response = gsub("mindfulresponse", "mindful response", response),
         response = gsub("cravingprevious", "craving previous", response),
         response = gsub("drinksnumber", "number of drinks", response),
         term = gsub(":", " x ", term),
         `b [95% CI]` = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
         path = ifelse(grepl("intervention", term) & response == "mindful response", "a1",
                ifelse(grepl("mindful", term) & response == "craving previous", "b1",
                ifelse(grepl("craving", term) & response == "number of drinks", "b2",
                ifelse(grepl("intervention", term) & response == "craving previous", "a2",
                ifelse(grepl("mindful", term) & response == "number of drinks", "b3",
                ifelse(grepl("intervention", term) & response == "number of drinks", "c'", ""))))))) %>%
  rename("outcome" = response) %>%
  select(outcome, path, term, `b [95% CI]`) %>%
  mutate(outcome = factor(outcome, levels = c("mindful response", "craving previous", "number of drinks"))) %>%
  arrange(outcome, path) %>%
  knitr::kable(digits = 2, format = "pandoc")
```

## indirect effect test {.tabset}
### intervention --> mindful responses --> alcohol consumption
```{r}
hypothesis(
  fit_brm_sequential,
  'b_mindfulresponse_active_weekon * b_drinksnumber_mindful_response +
  cor_pID__mindfulresponse_active_weekon__drinksnumber_mindful_response *
  sd_pID__mindfulresponse_active_weekon *
  sd_pID__drinksnumber_mindful_response
  = 0',
  class = NULL,
  seed  =  6523,
  alpha = .05
)
```

### intervention --> craving --> alcohol consumption
```{r}
hypothesis(
  fit_brm_sequential,
  'b_cravingprevious_active_weekon * b_drinksnumber_craving_previous +
  cor_pID__cravingprevious_active_weekon__drinksnumber_craving_previous *
  sd_pID__cravingprevious_active_weekon *
  sd_pID__drinksnumber_craving_previous
  = 0',
  class = NULL,
  seed  =  6523,
  alpha = .05
)
```

### intervention --> mindful responses --> craving --> alcohol consumption
Formula: E(a1,b1,b2) = E(a1)Cov(b1,b2) + E(b1)Cov(a1,b2) + E(b2)Cov(a1,b1) + E(a1)E(b1)E(b2)

```{r}
hypothesis(
  fit_brm_sequential,
  'b_mindfulresponse_active_weekon *
  cor_pID__cravingprevious_mindful_response__drinksnumber_craving_previous
  * sd_pID__cravingprevious_mindful_response * sd_pID__drinksnumber_craving_previous +
  
  b_cravingprevious_mindful_response *
  cor_pID__mindfulresponse_active_weekon__drinksnumber_craving_previous *
  sd_pID__mindfulresponse_active_weekon * sd_pID__drinksnumber_craving_previous +
  
  b_drinksnumber_craving_previous *
  cor_pID__mindfulresponse_active_weekon__cravingprevious_mindful_response *
  sd_pID__mindfulresponse_active_weekon * sd_pID__cravingprevious_mindful_response +
  
  b_mindfulresponse_active_weekon * b_cravingprevious_mindful_response *  b_drinksnumber_craving_previous
  
  = 0',
  class = NULL,
  seed  =  6523,
  alpha = .05
)
```

## summary
```{r}
summary(fit_brm_sequential, prob = .95)
```
